Performance Models for Data Transfers: A Case Study with Molecular Chemistry Kernels

### 2019-04-04 Article/article.tex
minor

	With the advent of multicore, and the use of accelerators, it is notoriously cumbersome to exploit the full capability of a machine. Indeed, there are several challenges come into picture. First, every architecture provides its own efficacy and interface. Therefore, a steep learning curve is required for programmers to take good utilization of all resources. Second, scheduling is a well known NP-Complete optimization problem, and hybrid and distributed resources make this problem harder (we refer ~\cite{webpagescheduling} for a survey on the complexity of scheduling problems and ~\cite{bleuse2015scheduling} for a recent survey in the case of hybrid nodes). Third, due to shared buses, parallel resources, it is hard to obtain a precise model based on prediction of computation and communication timings. Fourth, number of architectures has increased drastically in recent years, therefore it is almost impossible to develop hand tune optimized code for all these architectures. All these observations led to the development of different task based runtime systems. Among several runtimes, we may cite QUARK~\cite{YarKhan:2011:Quark:Manual} and PaRSEC~\cite{parsec} from from ICL, Univ. of Tennessee Knoxville (USA), StarPU~\cite{starpu} form INRIA Bordeaux (France), Legion~\cite{legion12} from Stanford Univ. (USA), StarSs~\cite{ompss} from Barcelona Supercomputing Center (Spain), KAPPI~\cite{kaapi} from INRIA Grenoble (France). All these runtime systems allow programmers to express their algorithms at the abstract level in the form of direct acyclic graphs (DAG), where vertices represent computations and edges represent dependencies among them. Sometimes some static information such as distance to exit (last) node as a priority or affinity of computation towards resources is also provided along with the DAG. Then runtime manages scheduling of computations and communications, data transfer among different memories, computation-communication overlap, load balance in the best possible way.

@@

	Certain applications such as dense linear algebra kernels have regular structure. Therefore, it is possible to associate priorities, based on the task graph structure, with computations, which runtime uses to make the execution efficient. In some applications, such as sparse linear algebra kernels, where programmers do not know the precise structure of the tasks graphs in advance,  tasks are added recursively based on certain sentinel constraints. For such applications, runtime sees a set of independent ready tasks and schedule them on different processing units. It is extremely important for runtimes to decide the order of data transfers for these scheduled computations such that overlap between computations and communications can be maximised. This is the main topic of this paper. We propose different strategies for runtimes such that it can maximize the overlap of computation and communication for kernels running on two different memory nodes. We prove that one of our proposed strategies is an optimal one when we do not have any restriction on the capacity of memory node. We also prove that order of communication on two memory nodes  with the objective of minimizing the makespan is  NP-Complete. We evaluate our strategies on a cluster of homogeneous nodes. But our approach is generic and easily adaptable to any system which operates on different memory spaces. Here are the important contributions of this paper.

@@

The outline of the article is the following. Section~\ref{sec:relatedWork} describes past work on the computations with limited memory and similar problems in the literature. In section~\ref{sec:theoreticalProof}, we present an algorithm to obtain the order of data transfers when there is not any memory capacity restriction and prove its optimality. Then, we also prove that in general data transfer problem is NP-complete. In Section~\ref{sec:heuristics}, we propose several heuristics and describe their favorable scenarios. We mainly consider three categories of heuristics: Static heuristics, dynamic heuristics and static heuristics with dynamic corrections. We evaluate our proposed strategies on two molecular chemistry kernels in Section~\ref{sec:exp}.  Our results show that static heuristics with dynamic corrections achieve good performance in most cases. We finally  propose conclusions and perspectives in Section~\ref{sec:conclusion}.

>>>>

	With the advent of multicore, and the use of accelerators, it is notoriously cumbersome to exploit the full capability of a machine. Indeed, there are several challenges come into picture. First, every architecture provides its own efficacy and interface. Therefore, a steep learning curve is required for programmers to take good utilization of all resources. Second, scheduling is a well known NP-Complete optimization problem, and hybrid and distributed resources make this problem harder (we refer ~\cite{webpagescheduling} for a survey on the complexity of scheduling problems and ~\cite{bleuse2015scheduling} for a recent survey in the case of hybrid nodes). Third, due to shared buses, parallel resources, it is hard to obtain a precise model based on prediction of computation and communication timings. Fourth, number of architectures has increased drastically in recent years, therefore it is almost impossible to develop hand tune optimized code for all these architectures. All these observations led to the development of different task based runtime systems. Among several runtimes, we may cite QUARK~\cite{YarKhan:2011:Quark:Manual} and PaRSEC~\cite{parsec} from from ICL, Univ. of Tennessee Knoxville (USA), StarPU~\cite{starpu} form INRIA Bordeaux (France), Legion~\cite{legion12} from Stanford Univ. (USA), StarSs~\cite{ompss} from Barcelona Supercomputing Center (Spain), KAPPI~\cite{kaapi} from INRIA Grenoble (France). All these runtime systems allow programmers to express their algorithms at the abstract level in the form of direct acyclic graphs (DAG), where vertices represent computations and edges represent dependencies among them. Sometimes some static information such as distance to exit (last) node as a priority or affinity of computation towards resources is also provided along with the DAG. Then runtime manages scheduling of computations and communications, data transfers among different memories, computation-communication overlap, load balance in the best possible way.

@@

	Certain applications such as dense linear algebra kernels have regular structure. Therefore, it is possible to associate priorities, based on the task graph structure, with computations, which runtime uses to make the execution efficient. In some applications, such as sparse linear algebra kernels, where programmers do not know the precise structure of the tasks graphs in advance,  tasks are added recursively based on certain sentinel constraints. For such applications, runtime sees a set of independent ready tasks and schedule them on different processing units. It is extremely important for runtimes to decide the order of data transfers for these scheduled computations such that overlap between computations and communications can be maximised. This is the main topic of this article. We propose different strategies for runtimes such that it can maximize the overlap of computation and communication for kernels running on two different memory nodes. We prove that one of our proposed strategies is an optimal one when we do not have any restriction on the capacity of memory node. We also prove that order of communications on two memory nodes  with the objective of minimizing the makespan is  NP-Complete. We evaluate our strategies on a cluster of homogeneous nodes. But our approach is generic and easily adaptable to any system which operates on different memory spaces. Here are the important contributions of this paper.

@@

The outline of the article is the following. Section~\ref{sec:relatedWork} describes past work on the computations with limited memory and similar problems in the literature. In section~\ref{sec:theoreticalProof}, we present an algorithm to obtain the order of data transfers when there is not any memory capacity restriction and prove its optimality. Then, we also prove that in general data transfer problem is NP-complete. In Section~\ref{sec:heuristics}, we propose several heuristics and describe their favorable scenarios. We mainly consider three categories of heuristics: static heuristics, dynamic heuristics and static heuristics with dynamic corrections. We evaluate our proposed strategies on two molecular chemistry kernels in Section~\ref{sec:exp}.  Our results show that static heuristics with dynamic corrections achieve good performance in most cases. We finally  propose conclusions and perspectives in Section~\ref{sec:conclusion}.
